---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = NULL
)
```

# alkahest

<!-- badges: start -->
[![R-CMD-check](https://github.com/tesselle/alkahest/workflows/R-CMD-check/badge.svg)](https://github.com/tesselle/alkahest/actions)
[![codecov](https://codecov.io/gh/tesselle/alkahest/branch/main/graph/badge.svg)](https://app.codecov.io/gh/tesselle/alkahest)
[![CodeFactor](https://www.codefactor.io/repository/github/tesselle/alkahest/badge/main)](https://www.codefactor.io/repository/github/tesselle/alkahest/overview/main)
[![Dependencies](https://tinyverse.netlify.com/badge/alkahest)](https://cran.r-project.org/package=alkahest)

[![r-universe](https://tesselle.r-universe.dev/badges/alkahest)](https://tesselle.r-universe.dev){.pkgdown-devel}
[![CRAN Version](http://www.r-pkg.org/badges/version/alkahest)](https://cran.r-project.org/package=alkahest){.pkgdown-release}
[![CRAN checks](https://cranchecks.info/badges/worst/alkahest)](https://cran.r-project.org/web/checks/check_results_alkahest.html){.pkgdown-release}
[![CRAN Downloads](http://cranlogs.r-pkg.org/badges/alkahest)](https://cran.r-project.org/package=alkahest){.pkgdown-release}

[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
<!-- badges: end -->

## Overview

**alkahest** is a lightweight, dependency-free toolbox for pre-processing XY data from experimental methods (i.e. any signal that can be measured along a continuous variable). It provides methods for baseline estimation and correction, smoothing, normalization, integration and peaks detection.

## Installation

You can install the released version of **alkahest** from [CRAN](https://CRAN.R-project.org) with:

```{r cran-installation, eval=FALSE}
install.packages("alkahest")
```

And the development version from [GitHub](https://github.com/) with:

```{r gh-installation, eval=FALSE}
# install.packages("remotes")
remotes::install_github("tesselle/alkahest")
```

## Usage

```{r packages}
## Load the package
library(alkahest)
```

**alkahest** expects the input data to be in the simplest form (a two-column matrix or data frame, a two-element list or two numeric vectors).

```{r baseline}
## gamma-ray spectrometry
data("BEGe")

## Subset from 3 to 200 keV
BEGe <- signal_select(BEGe, from = 3, to = 200)

## SNIP baseline
BEGe_snip <- baseline_snip(BEGe, LLS = FALSE, decreasing = FALSE, n = 100)

plot(BEGe, type = "l", xlab = "Energy (keV)", ylab = "Count")
lines(BEGe_snip, type = "l", col = "red")
```

```{r peaks}
## X-ray diffraction
data("XRD")

## Correct baseline
XRD_correct <- signal_correct(XRD, method = "SNIP")

## Find peaks
XRD_peaks <- peaks_find(XRD_correct, SNR = 3, m = 5)

plot(XRD_correct, type = "l", xlab = expression(2*theta), ylab = "Count")
lines(XRD_peaks, type = "p", pch = 16, col = "red")
```

```{r smooth, fig.width=5, fig.height=5, out.width='50%', fig.show='hold'}
## Simulate data
set.seed(12345)
x <- seq(-4, 4, length = 100)
y <- dnorm(x)
z <- y + rnorm(100, mean = 0, sd = 0.01) # Add some noise

## Plot raw data
plot(x, z, type = "l", xlab = "", ylab = "", main = "Raw data")
lines(x, y, type = "l", lty = 2, col = "red")

## Rectangular smoothing
unweighted <- smooth_rectangular(x, z, m = 3)
plot(unweighted, type = "l", xlab = "", ylab = "", main = "Rectangular smoothing") 
lines(x, y, type = "l", lty = 2, col = "red")

## Triangular smoothing
weighted <- smooth_triangular(x, z, m = 5)
plot(weighted, type = "l", xlab = "", ylab = "", main = "Triangular smoothing")
lines(x, y, type = "l", lty = 2, col = "red")

## Savitzky–Golay filter
savitzky <- smooth_savitzky(x, z, m = 21, p = 2)
plot(savitzky, type = "l", xlab = "", ylab = "", main = "Savitzky–Golay filter")
lines(x, y, type = "l", lty = 2, col = "red")

## Whittaker smoothing
whittaker <- smooth_whittaker(x, z, lambda = 1000, d = 3)
plot(whittaker, type = "l", xlab = "", ylab = "", main = "Whittaker smoothing")
lines(x, y, type = "l", lty = 2, col = "red")
```

## Contributing

Please note that the **alkahest** project is released with a [Contributor Code of Conduct](https://www.tesselle.org/conduct.html). By contributing to this project, you agree to abide by its terms.
