---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = NULL
)
```

# alkahest

<!-- badges: start -->
[![R-CMD-check](https://github.com/tesselle/alkahest/workflows/R-CMD-check/badge.svg)](https://github.com/tesselle/alkahest/actions)
[![codecov](https://codecov.io/gh/tesselle/alkahest/branch/master/graph/badge.svg)](https://app.codecov.io/gh/tesselle/alkahest)
[![CodeFactor](https://www.codefactor.io/repository/github/tesselle/alkahest/badge/main)](https://www.codefactor.io/repository/github/tesselle/alkahest/overview/main)

[![r-universe](https://tesselle.r-universe.dev/badges/alkahest)](https://tesselle.r-universe.dev){.pkgdown-devel}
[![CRAN Version](http://www.r-pkg.org/badges/version/alkahest)](https://cran.r-project.org/package=alkahest){.pkgdown-release}
[![CRAN checks](https://cranchecks.info/badges/worst/alkahest)](https://cran.r-project.org/web/checks/check_results_alkahest.html){.pkgdown-release}
[![CRAN Downloads](http://cranlogs.r-pkg.org/badges/alkahest)](https://cran.r-project.org/package=alkahest){.pkgdown-release}

[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
<!-- badges: end -->

## Overview

Pre-Processing XY Data from Experimental Methods. **alkahest** is a toolbox for pre-processing XY data from experimental methods (i.e. any signal that can be measured along a continuous variable). It provides methods for baseline correction, smoothing, normalization and integration.

This package depends only on the base R packages.

## Installation

You can install the released version of **alkahest** from [CRAN](https://CRAN.R-project.org) with:

```{r cran-installation, eval=FALSE}
install.packages("alkahest")
```

And the development version from [GitHub](https://github.com/) with:

```{r gh-installation, eval=FALSE}
# install.packages("remotes")
remotes::install_github("tesselle/alkahest")
```

## Usage

```{r packages}
## Load the package
library(alkahest)
```

**alkahest** expects the input data to be in the simplest form (a two-column matrix or data frame, a two-element list or two numeric vectors). With few exceptions, all functions return a list with two elements `x` and `y`.

```{r baseline}
## gamma-ray spectrometry
data("BEGe")

## Subset from 3 to 200 keV
BEGe <- signal_select(BEGe, from = 3, to = 200)

## SNIP baseline
BEGe_snip <- baseline_snip(BEGe, LLS = FALSE, decreasing = FALSE, n = 100)

plot(BEGe, type = "l", xlab = "Energy (keV)", ylab = "Count")
lines(BEGe_snip, type = "l", col = "red")
```

```{r smooth, fig.width=5, fig.height=5, out.width='50%', fig.show='hold'}
## gamma-ray spectrometry
data("LaBr")

## Subset from 1350 to 1650 keV
LaBr <- signal_select(LaBr, from = 1350, to = 1650)

## Plot raw spectrum
plot(LaBr, type = "l", xlab = "Energy (keV)", ylab = "Count", main = "Raw data")

## Rectangular smoothing
LaBr_unweighted <- smooth_rectangular(LaBr, m = 3)
plot(LaBr_unweighted, type = "l", xlab = "Energy (keV)", ylab = "Count",
     main = "Rectangular smoothing") 

## Triangular smoothing
LaBr_weighted <- smooth_triangular(LaBr, m = 5)
plot(LaBr_weighted, type = "l", xlab = "Energy (keV)", ylab = "Count",
     main = "Triangular smoothing")

## Savitzky–Golay filter
LaBr_savitzky <- smooth_savitzky(LaBr, m = 21, p = 2)
plot(LaBr_savitzky, type = "l", xlab = "Energy (keV)", ylab = "Count",
     main = "Savitzky–Golay filter")
```

## Contributing

Please note that the **alkahest** project is released with a [Contributor Code of Conduct](https://www.tesselle.org/conduct.html). By contributing to this project, you agree to abide by its terms.
